---
title: "Tree Data Analysis (ID issue fixed)"
author: "Tahmidul Islam"
date: "12/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, error=FALSE, message = FALSE, cache = FALSE, dpi = 300, fig.width = 14, fig.height = 6)
```

```{r}
library(dplyr)
library(ggplot2)
library(kableExtra)
library(tidyr)
```
# Tree Data

This dataset contains features of trees (height, width, etc.) from 3 sites: Brighton, Chase Stream and Lily Bay. There are 8 species in this sample.

Table: Data overview.
```{r}
tree <- read.csv("E:/R Project/BayesFDA/data/tree/tree.csv") 
head(tree) %>% kable()
glimpse(tree)
tree <- tree %>% dplyr::select(Site, ID.Code, Tree, Rep, Sp, Year, Height) %>% 
  filter(!is.na(Height)) %>% 
  mutate(id = paste0(Site, ID.Code, Rep, Tree), logHeight = log(Height))
head(tree) %>% kable()
```


Table: Distribution of site and species.
```{r}
sumDf <- tree %>% group_by(Site, Sp, id) %>% summarize (n = n()) %>% dplyr::select(Site, Sp)

sumDf %>% tally() %>% spread(Sp, n) %>% kable()
```

```{r}
ggplot(sumDf, aes(x = Site, fill = Sp)) + geom_bar(position="dodge", stat="count")
```

The measurements are collected few years apart: at 5, 10, 15, 16 and 27 years. 
```{r}
sumDf <- tree %>% group_by(Site, Year)  %>%  summarize(n = n())
sumDf %>% kable()
```

```{r}
ggplot(sumDf, aes(x = factor(Year), y = n, fill = Site)) + geom_col(position = 'dodge', stat = 'count')
```



```{r}
ggplot(data = tree, aes(x = Year, y = Height, col = Sp, group = id)) +
  geom_line(size = 1.5, alpha = .1) + facet_wrap(~Site)
```
```{r}
ggplot(data = tree, aes(x = Year, y = Height, col = Sp, group = id)) +
  geom_line(size = 1.5, alpha = .1)
```

## GP model: Site Specific Mean Height Estimation

### Parameter estimates 

```{=latex}
\begin{table}[ht]
\centering
\begin{tabular}{c|cc|cc|cc|c}
  \hline
 & \multicolumn{2}{|c|}{Mean} & \multicolumn{2}{|c|}{Subject} & \multicolumn{2}{|c|}{Site} & \\
 \hline
 & $l^2$ & $\sigma^2$ & $l^2$ & $\sigma^2$ & $l^2$ & $\sigma^2$ & $\sigma^2$ \\ 
  \hline
BS & 0.69 & 1.79 & 0.17 & 0.06 & 0.14 & 0.00 & 0.03 \\ 
  EL & 1.06 & 4.80 & 0.20 & 0.05 & 0.08 & 0.01 & 0.02 \\ 
  HL & 2.09 & 5.96 & 0.27 & 0.07 & 11.55 & 0.01 & 0.02 \\ 
  JL & 1.69 & 12.24 & 0.21 & 0.03 & 0.19 & 0.02 & 0.01 \\ 
  JP & 4.36 & 14.42 & 0.18 & 0.02 & 0.10 & 0.01 & 0.02 \\ 
  RP & 0.86 & 2.50 & 0.16 & 0.05 & 0.01 & 0.00 & 0.01 \\ 
  TL & 0.28 & 1.60 & 0.20 & 0.04 & 0.01 & 0.00 & 0.02 \\ 
  WS & 0.67 & 2.08 & 0.20 & 0.16 & 24.91 & 0.00 & 0.01 \\ 
   \hline
\end{tabular}
\end{table}
```

### Plots of site specific estimation

```{r site, echo = FALSE, message=FALSE, fig.align='center', out.width='.9\\linewidth'}
knitr::include_graphics("../plot/tree/bs_site.pdf")
knitr::include_graphics("../plot/tree/el_site.pdf")
knitr::include_graphics("../plot/tree/hl_site.pdf")
knitr::include_graphics("../plot/tree/jl_site.pdf")
knitr::include_graphics("../plot/tree/jp_site.pdf")
knitr::include_graphics("../plot/tree/rp_site.pdf")
knitr::include_graphics("../plot/tree/tl_site.pdf")
knitr::include_graphics("../plot/tree/ws_site.pdf")
```


# Classification

We further predict the tree species based on their growth (height) over time. The classification result is summarized in the following table in terms of a confusion matrix. The overall accuracy for prediciting the tree species is 46.14% which is not very high. The table also shows the percentage of trees correctly classified given the species. It appears that TL trees are classified best (71% classified as TL) whereas the larches are somewhat difficult to classify (EL: 27%, JL: 24%). However, interestingly the trees are 

```{=latex}
\begin{table}[ht]
\small
\centering
\caption{Confusion matrix for predicting tree species. The percentages in the braces give the accuracy of predicting the species given a species.}
\begin{tabular}{cccccccccc} 
\toprule
     & \multicolumn{8}{c}{Predicted}                 & Total  \\
True & BS & EL & HL & JL & JP & RP & TL & WS  &        \\
  BS & 149 (54.38)&   0 &   0 &   0 &  19 &  11 &   0 &  95 &  274 \\ 
  EL &  15 & 258 (27.27) & 377 & 156 &   9 &   8 & 105 &  18 &  946 \\ 
  HL &   7 &  37 & 190 (66.20) &  17 &   6 &   3 &  24 &   3 &  287 \\ 
  JL &   2 & 149 & 118 & 116 (24.47) &   4 &  13 &  63 &   9 &  474 \\ 
  JP &  43 &   0 &   0 &   4 & 160 (55.56)&  25 &  40 &  16 &  288 \\ 
  RP &  28 &   0 &   0 &   7 &  66 & 138 (46.00)&   4 &  57 &  300 \\ 
  TL &  15 &  23 &   7 &  40 &  41 &  12 & 379 (71.11)&  16 &  533 \\ 
  WS &  67 &   0 &   0 &   0 &  11 &  45 &   0 & 182 (59.67)&  305 \\ 
\bottomrule
\end{tabular}
\end{table}
```


We can aggregate the prediction results to compute the accuracy when trying to predict a tree's genus rather than the species. In this case, the overall accuracy jumps to 86.32%. Therefore, our model performs much better at predicting a tree's genus than the exact species.


```{=latex}
\begin{table}[ht]
\centering
\caption{Confusion matrix for predicting tree genus The percentages in the braces give the accuracy of predicting the genus given a species.}
\begin{tabular}{cccccccccc} 
\toprule
     & \multicolumn{3}{c}{Predicted} & Total  \\
    True & Spruce  & Larch  & Pine  &        \\
  BS & 244 (89.05\%) &  0   &  30 &   274 \\ 
  EL &  33 &  896 (94.71\%) &  17 &  946 \\ 
  HL &  10 &  268 (93.38\%) &  9  &  287 \\ 
  JL &  11 &  446 (94.09\%) &  17 &  474 \\ 
  JP &  59 &  44  & 185 (64.24\%) &  288 \\ 
  RP &  85 &  11  & 204 (68.00\%) &  300 \\ 
  TL &  31 &  449 (84.24\%) & 53  &  533 \\ 
  WS &  249 (81.64\%) &   0 & 56  &  305 \\ 

\bottomrule
\end{tabular}
\end{table}
```

